Transparency of COVID-19 medical literature

### 2022-03-06 protocol_covid_transparency.qmd
Update protocol_covid_transparency.qmd
protocol addition uribe plus one reference

---
title: "Transparency of COVID-19 medical literature"
author:
  - Ahmad Sofi-Mahmudi^[Seqiz Health Network, Kurdistan University of Medical Sciences, Seqiz, Iran.; a.sofimahmudi@gmail.com]
  - Eero Raittio*^[Institute of Dentistry, University of Eastern Finland, Kuopio, Finland.; eero.raittio@uef.fi]
  - Sergio E. Uribe^[Department of Conservative Dentistry and Oral Health & Bioinformatics Research Unit, Riga Stradins University, Riga, Latvia.; sergio.uribe@rsu.lv]
format: html
editor: visual
bibliography: references.bib
csl: vancouver.csl
---

**Corresponding author:** Eero Raittio; Address: ; Telephone: ; Email: eero.raittio\@uef.fi.

@@

**Objective and aim:** We aim to assess the use of transparency indicators (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access COVID-19-related articles published in medical journals available from the Europe PubMed Central (PMC) database. We will map transparency in medical articles across different journals, publication years, publishers, specialties and journal impact factors.

**Methods:** We will search and export all open access articles from PubMed-indexed journals available in the Europe PMC database. We will detect transparency indicators with a validated and automated tool developed to extract the indicators from the downloaded articles. Basic journal- and article-related information will be retrieved from the PMC database. We will use R for the searches (europepmc package), data handling, analysis (rtransparent package) and reporting (ggplot2 package).

@@

Our aim is to assess the uptake of transparent practices (data sharing, code sharing, conflict of interest disclosures, funding disclosures, statements of protocol registration) from open access full-text COVID-19-related articles published in medical journals available from the Europe PMC. We map transparency across publication years, publishers and journals.

@@

The COVID-19 query is available in the *covid_keywords.csv* file.

@@

We aim to describe the use of five transparency indicators:

@@

@@

The study sample will be restricted to open access articles in the Europe PMC database which may not correspond to all COVID-19-related studies published in medical journals. However, as the majority of COVID-19-related papers are open access, this will not diminish the strength of our interpretations considerably.

>>>>

---
title: "Transparency of COVID-19 medical literature"

author:
  - Ahmad Sofi-Mahmudi^[Seqiz Health Network, Kurdistan University of Medical Sciences, Seqiz, Iran.; a.sofimahmudi@gmail.com]
  - Eero Raittio*^[Institute of Dentistry, University of Eastern Finland, Kuopio, Finland.; eero.raittio@uef.fi]
  - Sergio E. Uribe^[Department of Conservative Dentistry and Oral Health & Bioinformatics Research Unit, Riga Stradins University, Riga, Latvia.; sergio.uribe@rsu.lv] ^[School of Dentistry, Universidad Austral de Chile, Valdivia, Chile]

format: html
editor: visual
bibliography: references.bib
csl: vancouver.csl
---

**Corresponding author:** Eero Raittio; **Address**: ; Telephone: ; Email: eero.raittio\@uef.fi.

@@

**Objective and aim:** The recent pandemic has shown how crucial it is to have access to quality medical information and data that allows people to have confidence in healthcare decisions. One of these indicators is transparency, i.e., the availability of original data supporting research results. We aim to assess the use of transparency indicators (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access COVID-19-related articles published in medical journals available from the Europe PubMed Central (PMC) database. We will map transparency in medical articles across different journals, publication years, publishers, specialties and journal impact factors.

**Methods:** We will search and export all open access articles from PubMed-indexed journals available in the Europe PMC database published between January 2020 and December 2021. We will detect transparency indicators with a validated and automated tool developed to extract the indicators from the downloaded articles. Basic journal- and article-related information will be retrieved from the PMC database. We will use R for the searches (europepmc package), data handling, analysis (rtransparent package) and reporting (ggplot2 package).

@@

Our objective is to evaluate the transparency indicators (data sharing, code, declaration of interest, funding, availability of research protocol) from open access full-text COVID-19-related articles published in medical journals available from the Europe PMC. We map transparency across publication years, publishers and journals.

@@

The full COVID-19 query is available in the *covid_keywords.csv* file.

@@

We aim to describe adherence to five transparency practices in the articles:

@@

We will also calculate the proportion of articles available as open access via the Europe PMC from the total number of COVID-19-related detected in the database.

@@

The study sample will be restricted to open access articles in the Europe PMC database which may not correspond to all COVID-19-related studies published in medical journals. However, as the majority of COVID-19-related papers are open access[@Kiley2020-ek], this will not diminish the strength of our interpretations considerably.
